---
title: "covid 19"
author: "Isai Bercian"
date: "3/18/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$:Covid 19 deaths in Korea, Jan-May 2020 are unrelated to sex.
$H_A$:there is a relationship between deaths in Korea Jan-May 2020 and sex of the patient.

Table of patients:
```{r}
kRdata$Sex = as.factor(kRdata$Sex)
kRdata$result= as.factor(kRdata$result)

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))
levels(kRdata$Sex)

```
According to the table of patients dying is not independent of sex because only 25 females died compared to 44 men that died.

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


According to the barchart around 1800 females are alive compared to around 1300 males that are alive. On the percent barchart it shows around 65 males died from covid 19 in Korea while only around 35 females died from covid 19 in Korea.
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that 98.7% of females are alive and 1.4% of them died from covid 19 in Korea from Jan-May. Also 97% of males are alive and 3% of them died from covid 19 in Korea from Jan-May.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



## Conclusions

In conclusion we reject the null hypothesis because the p value is 0.001 and the odds of a male dying by covid 19 in Korea are 2.3 times greater then that of a female dying from covid 19 in Korea. Also because the confidence interval does not contain the number 1.